---
title: "03_PCoA-permanova"
author: "Sara Correa Garcia"
format: html
---

********************************************************
# LOAD ENV
# ********************************************************

``` {r}
rm(list = ls())
source(here::here("source", "libraries.R"))
load(file = here::here("RData", "clean_Data_rarefied.RData"))
```

********************************************************
# DISSIMILARITY MATRIX
# ********************************************************

## Calculate relative abundance from rarefied data
```{r}
com_rel <- (com_sorted/rowSums(com_sorted)) #Normalization by relative abundance
sanity.check <- rowSums(com_rel) #Sanity check
rownames(meta_sorted) == rownames(com_rel) # All true
```

## Calculate Bray Curtis dissimilarity 
from relative abundance on rarefied data
```{r}
bray_dist <- vegdist(com_rel, method = "bray")
```

#********************************************************
# PERMANOVA
# ********************************************************
## Permanova procedure
```{r}
set.seed(27)
perm <- adonis2(bray_dist~Water_type*Sample_type*Time, data = meta_sorted, permutations = 999)
as.data.frame(perm)
#write.csv(file = here::here("output", "tables", "permanova_full_model_16S-water.csv"), perm)
```

## Betadisper 
(multivariate analogue of Levene's test for homogeneity of variances) - Marti Anderson's PERMDISP2.
### Calculate multivariate dispersions
```{r}
dispersion <- betadisper(bray_dist, meta_sorted$treatment, sqrt.dist = TRUE, type = "centroid", bias.adjust = FALSE)
```

### perform test
```{r}
anova(dispersion) 
```

### Permutation test for F
```{r}
permutest(dispersion, pairwise = TRUE, permutations = 999)
```

### Plot the groups and distance to centroids on the first two PCoA axes
```{r}
plot(dispersion)
plot(dispersion, ellipse = TRUE, hull = FALSE) # 90% data ellipse
```

### Boxplot of the distances to centroid for each group
```{r}
boxplot(dispersion)
```

#********************************************************
# INDICATOR SPECIES
# ********************************************************
## Indicspecies
```{r}
library(indicspecies)
inval <- indicspecies::multipatt(com_sorted, meta_sorted$treatment, control = how(nperm = 999))
summary(inval, alpha = 0.001)
dim(com_sorted)
```


#********************************************************
# PCoA
# ********************************************************

## PCoA calculation
```{r}
pcoa <- cmdscale(sqrt(bray_dist), eig = TRUE)
```

##Calculate % variability 
explained by eigenvalues 1 and 2.
```{r}
explainedvar1 <- (pcoa$eig[1]/sum(pcoa$eig)*100)#Axis1 explains 11.06879% of the variance
explainedvar1
explainedvar2 <- (pcoa$eig[2]/sum(pcoa$eig)*100)#Axis2 explains 9.358371% of the variance
explainedvar2
sumeigen <- (explainedvar1 + explainedvar2)# 20.42716%
sumeigen
```

Axis 1 explains `r round(explainedvar1, digits=2)`% of the total variability, whereas the second axis explains `r round(explainedvar2, digits=2)`%. together, axis 1 and 2 explain `r round(sumeigen, digits = 2)`% of the variance.

## Create coordinates
for axis 1 and 2 for PCoA from eigenvalues
```{r}
pcoa_points <- data.frame(pcoa$points) # create a column with the coordinates for each sample
colnames(pcoa_points) = c("Axis1", "Axis2") # rename coordinate columns
```

## Bind axes coords to meta
need to be in the same object, ordered community and ordered metadata
```{r}
pcoa_points_bind <- cbind(pcoa_points, meta_sorted) #
str(pcoa_points_bind)
```

## PCoA PLOT

```{r}
pcoa_points_bind$Time <- factor(pcoa_points_bind$Time, levels = c("D-0", "D8", "D28", "D42", "D84"))
pcoa_points_bind$treatment <- factor(pcoa_points_bind$treatment, labels = c("Lab water", "OSPW", "OSPW + Carex"))
pcoa_plot <- ggplot(data = pcoa_points_bind, aes(x = Axis1, y = Axis2)) +
        geom_vline(xintercept = 0, linetype = 'twodash', col = 'grey') +
        geom_hline(yintercept = 0, linetype = 'twodash', col = 'grey') +
        geom_point(aes(shape = Time, color = treatment, size = 5)) + 
        #scale_shape_manual(values = c(16,17,1)) +
        scale_color_manual(values = c("#009ACD", "#EEC900", "#787878")) + 
        xlab(glue("PCo 1 = {round(explainedvar1, 2)} %")) + 
        ylab(glue("PCo 2 = {round(explainedvar2, 2)} %")) + 
        ggtitle("PCoA with BC Dissimilarity of the aquatic bacterial community") +
        scale_size(guide="none") + 
    #+  # to hide the part of the legend corresponding to size
        theme_classic() +
        theme_linedraw() +
        theme(axis.title.x = element_text(size = 24, face = "bold"),
              axis.text.x = element_text(size = 20),
              panel.grid = element_blank(),
              #axis.ticks.x = element_blank(),
              axis.title.y = element_text(size = 24, face = "bold"),
              axis.text.y = element_text(size = 20),
              legend.position = "right",
              legend.text = element_text(size = 18),
              legend.title = element_blank(),
              #axis.line = element_line(linewidth = 1),
              plot.title = element_text(hjust = 0.5, face = "bold")) +
        guides(colour = guide_legend(override.aes = list(size=5))) +
        guides(shape = guide_legend(override.aes = list(size=5))) 
pcoa_plot

ggsave(file = here("output","figures", "PCoA_BC_rarefied.pdf"), pcoa_plot, height = 8, width = 10, units = "in")
ggsave(file = here("output","figures", "PCoA_BC_rarefied.eps"), pcoa_plot, height = 8, width = 10, units = "in")
```


#********************************************************
# END OF SCRIPT
# ********************************************************

##Palettes
```{r}
earthy_pal <- c("#E8B89A", "#BBB7AC", "#506A67", "#BF795F","#506A67", "#434341", "#E2DFD8")
c( "#3b528b", "#EEC900", "#1F1F1F", "#FF7F00", "#9400D3","#00BFFF",)
c("#fde725", "#5ec962","#21918c", "#3b528b", "#440154")
c("#787878", "#9C9C9C", "#BABABA", "#C4C4C4", "#E5E5E5", "#F2F2F2")
c("#104E8B", "#1874CD", "#1E90FF", "#009ACD", "#63B8FF", "#4F94CD", "#87CEFF", "#4876FF", "#C6E2FF", "#3A5FCD", "#87CEFA", "#BFEFFF")
```

